---
title: Hire Me (as a Data Scientist!), Part II
author: ~
date: '2018-01-15'
slug: hire-me-as-a-data-scientist-part-ii
categories: [R, 10 Minute Reads]
tags: [R, Data Science, Beer]
---


Continuing on from my earlier post, I'm now looking to tackle the question:

> If you had to pick 3 beers to recommend using only this data, which would you pick?

This is a pretty open ended question, which is kind of fun.
I also don't really have a ton of experience (yet!) in recommendation systems, though I have done a little reading here or there on it.

My goals in coming up with three beers to recommend were to:

1. Try to find the most popular beer among super users of the website
2. Find a [bizzaro](https://youtu.be/IcjSDZNbOs0?t=31s) beer that matched the profile of my first beer, but lives in the [long tail](https://books.google.com/books?hl=en&lr=&id=DTeZAAAAQBAJ&oi=fnd&pg=PT6&dq=anderson+2006+long+tail&ots=MpaGpMbdfD&sig=25QPk_RCCNU2yFoo9nsU0hrt0sc#v=onepage&q=anderson%202006%20long%20tail&f=false) of the ratings distribution
3. Find the best Beer sans Booze (Highest Rating with lowest ABV)

So let's begin!
Here's how I went about tackling this question.

## Popular with Super Users

```{r }
#=====================================================================================#
# Following suit of the last post... 
#=====================================================================================#
# Library
library(ggplot2)
library(data.table)
library(stringr)
#=====================================================================================#
beer <- fread("data/beer_reviews.csv")
beer.complete <- beer[complete.cases(beer)]
#=====================================================================================#
```

Having more experience in experimental settings, one of the first things I needed to get used to when I started working with non-psychology datasets was the lack of complete sets in what felt like almost everything.
Whereas in the [lab](https://musiccog.lsu.edu/) we spend lots of time trying to design balanced studies that hopefully don't violate the litany of assumptions that classic null hypothesis significance testing demands, my first few attempts at analyzing large amounts of data made me realize it's almost risible to think that you're going to have even, independent data, ever. 
This dataset is no different.

Of all of the unique users on the site, most of them have done only a couple of reviews, but some have essentially made a job out of this.
Looking at the distribution of reviews, this is quite clear. 

```{r}
review.counts <- beer[, .(.N), by = review_profilename][order(-N)]
review.counts 
hist(beer[, .(.N), by = review_profilename][order(-N)]$N, 
     breaks = 200,
     xlab = "Number of Reviews",
     main = "Distribution of Reviews Per User")

```

This is pretty important when it comes to modeling the data (discussed in Part III), and not being fully aware of where your ratings are coming from could put the quality of your models at serious risk.

So looking at this dataset, I wondered if there were any sort of implicit assumptions I could make about this data that might be able to help me find a good beer.
One assumption that I didn’t think was too wild was that a sample of this population that had gone out of its way to rate over 500 beers was probably more of a beer expert than those who have only done a couple of reviews on the site.

One thing I wanted to check was: of all the 1.5 million reviews, where were they coming from? Were there enough reviews among the super users that I could use? 
And what made someone a super user? 
I could have been a bit more scientific, setting an _a priori_ threshold, but for this I kind of just looked at that chart above, spit balled thinking 500 might be a good number to check, and then went to see how much of the data would be accounted if I put my threshold there. 
I lucked out and got about half of it.


```{r}
sum(review.counts$N) # Number of Total Reviews 
sum(review.counts[ N > 500]$N) # Number of Reviews from Super Users

731066/1586614 # Percent of Total Reviews from 500+ Super Users

super.users <- review.counts[ N >  500] # I can settle for .75 Million Reviews
```

So now I had a list of the users who had completed over 500 reviews and made up 46% of our entire data. 
I could use this new table I had made to index through our dataset of all the reviews that I had (that have their ABV ratings!) so I was then only dealing with these higher quality reviewers.

```{r}
super.reviews <- super.users[beer.complete, on = "review_profilename", nomatch=0]
```

As I continued to chop down the dataset (since this was a very exploratory process compared with cleaning up an experiment), it was important to do **quality assessment** steps. 
One thing worth checking here was to see if I was actually dealing with beer omnivores in our super users and make sure that all different types of beers were being represented in our smaller subset. 
This was done by just looking at the number of rows between the original dataset and our super user table.

```{r}
super.reviews[, .(beer_styles = unique(beer_style))]
beer[, .(beer_styles = unique(beer_style))]
```

Luckily they were the same.
If I were to really do some more work on this dataset, I would also want to check things such as how many of the beers had each super user tried?
Were there IPA experts in the group?
If yes, should their opinions be taken more seriously if I had questions about IPA recommendations in the future?
But for now, I just set out to see what the highest rated beer among all the super users of this dataset was.

In order to answer that question, I had to find out which beer in specific had the highest mean rating. 
The dataset ‘as is’ comes with a ```beer_id``` unique ID, but the data downloaded as is does not give us a key to this, so I had to make it myself. 
This was accomplished by just pasting together the brewery’s name, along with the beer name, and style into a new variable.

As another **quality assurance** step, it was worth checking to see if this recreated the unique ID variable, which it didn’t do exactly... but it was pretty close. 
I would chalk that up to some sort of encoding error.

```{r}
super.reviews[, beer_name_unique := paste(brewery_name,beer_name, beer_style)]

length(unique(super.reviews$beer_beerid))
length(unique(super.reviews$beer_name_unique))
42703/42805 # Pretty close

super.reviews.popular <- super.reviews[, .(most_reviewed_beers = .N), 
                                       by = beer_name_unique][order(-most_reviewed_beers)]

hist(super.reviews.popular$most_reviewed_beers,
     main = "Distribution of Number of Ratings by Super Users",
     xlab = "Number of Reviews each Beer Recieves",
     breaks = 200)
```

Again I saw this was clearly **not** anything resembling a repeated measures experiment and not all beers were rated equally. 

Continuing in the same fashion above, I just grabbed the top 100 beers of our super users and merged that on to our table from earlier that had all of the ratings from our super users. 
Then from that table, I took the average of the overall rating and looked at our top ten beers.

```{r}
super.reviews.popular.100 <- super.reviews[, .(most_reviewed_beers = .N), by = beer_name_unique][order(-most_reviewed_beers)][1:100]

super.reviews.cream.of.crop <- super.reviews.popular.100[super.reviews, 
                                                         on = "beer_name_unique", 
                                                         nomatch=0]

super.reviews.cream.of.crop[, .(mean_review_overall = mean(review_overall)), by = beer_name_unique][order(-mean_review_overall)][1:10]
```

And we have our winner!
It's [Pliny The Elder](https://www.beeradvocate.com/beer/profile/863/7971/) from Russian River Brewing Company as my first beer recommendation!

## Bizzaro Beer

Now Pliny The Elder seemed to be a pretty popular beer. 
But what if I was trying to sketch out some ideas about what other beers I could recommend to beer lovers who like Pliny The Elder? 
It needed to somewhat “look like” the target beer, but have way less reviews.

Playing with some of the fringe data here, I wanted to be careful not to again pick a beer with only one or two ratings on it. 
My rationale was coming from assuming there is some sort of true “population mean” for this beer and having a beer with too little reviews will not approximate the mean correctly.

```{r}
## Make Unique Beer Label for Larger Dataset
beer[, beer_name_unique := paste(brewery_name,beer_name, beer_style)]

## Count Number of Reviews Each Beer Has  
number.of.reviews <- beer[, .(NumberOfReviews = .N), by = beer_name_unique][order(-NumberOfReviews)]
 
## Only get beers with over 30 reviews
reliable.beers.list <- number.of.reviews[ NumberOfReviews >= 30 ]

## Join that to our big 'beer' dataset only matching beers with over 30 reviews
beer.reliable <- reliable.beers.list[beer, on = "beer_name_unique", nomatch=0]
```

With our dataset chiseled down to only 'reliable' beers, I needed to find a way to get some sort of profile of each of the beers.
While my first instinct was to do some sort of data reductive type thing like a PCA on our continuous variables and use distances from certain scores as metrics of similarity ([which I have done before](https://musiccog.lsu.edu/davidjohnbaker/papers/Baker_Trahan_Mullensiefen_ProceedingsPaper.pdf) and it ended up actually being the inspiration for a tool currently used by [Soundout](https://www.soundout.com/brandmatch)!), doing that on so few predictors seemed [extra](https://www.urbandictionary.com/define.php?term=extra).

So instead, I figured why not just assume that there is some sort of wiggle room in my hastily made recommendation system and just match first on the overall review, then if there are some close contenders, look for matches on other metrics?

The next bit of code creates a table of the metrics I am interested in, gets beers that have over 30 reviews, but less than 100, and also creates a vector so I can pull out all of the IPAs on my less reviewed beers table.
I then joined the tables for my candidates. 
```{r}
# # Get metrics used for distance calculations
beer.metrics <- beer.reliable[, .(mean_review_overall = mean(review_overall),
                                  mean_review_aroma = mean(review_overall),
                                   mean_review_appearance = mean(review_appearance),
                                   mean_review_palate = mean(review_palate),
                                   mean_reviw_taste = mean(review_taste),
                                   sd_review_overall = sd(review_overall),
                                   sd_review_aroma = sd(review_overall),
                                   sd_review_appearance = sd(review_appearance),
                                   sd_review_palate = sd(review_palate),
                                   sd_review_taste = sd(review_taste)),
                               by = beer_name_unique]

## Get only IPAs with less than 100 reviews
less.reviewed.beers <- number.of.reviews[NumberOfReviews <= 100 & NumberOfReviews >= 30]
## Make vector to help find IPAs
find.IPA <- str_detect(string = less.reviewed.beers$beer_name_unique, pattern = "Imperial IPA")
bizzaro.candidates <- less.reviewed.beers[find.IPA]

#Create Table
bizzaro.candidates.metrics <- bizzaro.candidates[beer.metrics, on = "beer_name_unique", nomatch=0]
```

Of these less reviewed beers, I now needed to find the one that was “closest” on the few dimensions I had to work with. 
The simplest way to do this would be to just subtract our target beer (Pliny The Elder), from every other beer in our interested list, then check out the results.

```{r}
## Get metrics for our target beer
rrbcpteadii.metrics <- beer[beer_name_unique == "Russian River Brewing Company Pliny The Elder American Double / Imperial IPA",
                           .(mean_review_overall = mean(review_overall),
                             mean_review_aroma = mean(review_overall),
                             mean_review_appearance = mean(review_appearance),
                             mean_review_palate = mean(review_palate),
                             mean_reviw_taste = mean(review_taste),
                             sd_review_overall = sd(review_overall),
                             sd_review_aroma = sd(review_overall),
                             sd_review_appearance = sd(review_appearance),
                             sd_review_palate = sd(review_palate),
                             sd_review_taste = sd(review_taste))]
## Create vector for looping over
key.vector <- as.vector(rrbcpteadii.metrics)
## Pull off the tags of our search
search.vector <- bizzaro.candidates.metrics[, -c(1,2)]

## Sanity check that what we are going to subtract has same names
names(key.vector)
names(search.vector)

## And that the apply function I am going to run is doing what I think it will
search.vector[1]- key.vector

## Run apply function
ipa.distances <- apply(search.vector, 1, function(x) x - key.vector)
ipa.distances.dt <- data.table(do.call(rbind.data.frame,ipa.distances))
## Combine this back with vector with names
bizzaro.candidates.distances <- cbind(bizzaro.candidates.metrics, ipa.distances.dt)
## Sort our data by overall and see if we have a good match!
bizzaro.candidates.distances[order(-mean_review_overall)]
```
Of course if I were building a real recommendation machine I could start talking about what factors are more important for what users and what factors are more predictive than others, but this seems like an OK enough solution to at least have completed my _a priori_ goal.

Based on this solution, it looks like I will have to find myself a bottle of [Pliny The Elder](https://www.beeradvocate.com/beer/profile/863/7971/) and the [Hill Farmstead Brewery Galaxy Imperial Single Hop IPA American Double / Imperial IPA](https://www.beeradvocate.com/beer/profile/22511/67760/) and do some of my own empirical work to see if this was a good idea. 

## Beer sans Booze 

The last beer that I think I wanted to recommend would be one that tastes great, but does not have a lot of alcohol in it.
The reason this question kind of interests me is because if we are _really_ going to talk about how tasty a beer is, it would be nice to be able to factor out of the equation how drunk we are actually getting from it.

I can see first of all IF this relationship exists if we look at the mean overall rating of a beer as a function of its ABV content.

```{r}
beer.complete[, beer_name_unique := paste(brewery_name,beer_name, beer_style) ]

# ABVs and Mean Scores
abv.vs.mean <- beer.complete[, .(Abv = mean(beer_abv), MeanOverall = mean(review_overall)), by = beer_name_unique]

ggplot(abv.vs.mean[Abv < 20], aes(x = Abv, y = MeanOverall)) + 
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  geom_smooth(method = "lm", formula = y ~ poly(x,2), color = "orange") +
  labs(title = "Rating as Function of ABV (Beers with than 20% ABV)",
       x = "ABV Content",
       y = "Mean Overall Rating")
```

Surprisingly, when I ran some quick and dirty regression models (that yes, I know violate tons of assumptions) I saw that only a very small amount of variance was being explained by its ABV. 
Note that although the models were significant, the R squared values hovered around 3-5%!

```{r}
# "The More Booze The Better" Model
abv.linear <- lm(MeanOverall ~ Abv, data = abv.vs.mean[Abv < 20]) 
# The "Diminishing Returns Model "
abv.poly <- lm(MeanOverall ~ poly(Abv,2), data = abv.vs.mean[Abv < 20])
# The "Dissapointing Amount of Variance Explained Summaries"
summary(abv.linear)
summary(abv.poly)
```

This actually surprised me and might be worth looking into at a deeper level another time, but for now I want to keep going on and find a beer knowing that how much booze is in it doesn't really affect how good it is.

So let's take one final dive into the dataset, grab only our quality reviews then plot a subset of our data so I can see beers that have a very high overall rating with a very small amount of booze in them. 

```{r}
# Quality Assurance Step
reliable.and.abv.beers <- reliable.beers.list[beer.complete, on = "beer_name_unique", nomatch=0]

## Get mean ratings and keep ABV (which won't change if I average it)
dd.beers <- reliable.and.abv.beers[, .(mean_overall = mean(review_overall), abv = mean(beer_abv)), by = "beer_name_unique"]

# Only Beers that Fit Our Criterion
dd.beers.2 <- dd.beers[mean_overall > 4.6 & abv < 10]

# Plot It!
ggplot(dd.beers.2, aes(x = abv, y = mean_overall, label = beer_name_unique, color = beer_name_unique)) +
  geom_point() +
  geom_text(aes(label=beer_name_unique),hjust=-.01, vjust=0) +
  labs(title = "High Quality Beers with Low ABV",
       x = "ABV",
       y = "Mean Overall Rating") + theme(legend.position = "none") +
  xlim(0, 20) +
  scale_y_continuous(breaks = c(seq(4.6,5,.1)), limits = c(4.6,4.85))
```

These are all OK choices (most of the beers are still above 5% ABV), but we do have one beer clocking in at 2.0% ABV giving us our final beer recommendation -- the [Southampton Publick House Southampton Berliner Weisse Berliner Weissbier](https://www.beeradvocate.com/beer/profile/1628/8626/)!

## Summary

After all of this, I know have three beers to check out.
[Pliny The Elder](https://www.beeradvocate.com/beer/profile/863/7971/) is our winner for the top rated beer among our Super Users of the site, the [Hill Farmstead Brewery Galaxy Imperial Single Hop IPA American Double / Imperial IPA](https://www.beeradvocate.com/beer/profile/22511/67760/) is a beer to maybe follow up on from our first choice, and then lastly we have the the [Southampton Publick House Southampton Berliner Weisse Berliner Weissbier](https://www.beeradvocate.com/beer/profile/1628/8626/) which supposedly tastes great, despite its lack of alcohol content. 
