---
title: Hire Me (as a Data Scientist!), Part IV
author: ~
date: '2018-01-29'
slug: hire-me-as-a-data-scientist-part-iv
categories: [R, 10 Minute Reads]
tags: [R, data science, beer]
---

Since this is my last post in the beer review series, I'll keep this short and sweet in terms of analysis.
Having done all of this, I do have a few reflections I would like to share after doing the [One-Size-Fits-All-Data-Science Interview](https://www.linkedin.com/pulse/how-hire-test-data-skills-one-size-fits-all-interview-tanya-cashorali/) that I have included at the end.

Our final question to answer is: 

> Lastly, if I typically enjoy a beer due to its aroma and appearance, which beer style should I try?

This is a pretty broad question and should be able to be answered without many pitfalls, so let's get started. 

```{r, eval=FALSE}
library(ggplot2)
library(data.table)
beer <- fread("data/beer_reviews.csv") 
```

From a bird's eye view, it seems like the most sensible thing to do would be to look at our data
from the highest level, then just zoom in until we have the level of granularity we feel answers
the question well.
Let's first average all the beer styles to get a rough estimate of how a beer style fairs
on the variables we are interested in, and additionally see how much variability is associated
with that measurement.

```{r}
sem <-function(x) {sd(x)/sqrt(length(x))}

question4.means <- beer[, .(mean_aroma = mean(review_aroma), mean_appearance = mean(review_appearance), mean_overall = mean(review_overall),
                            sem_aroma = sem(review_aroma), sem_appearance = sem(review_appearance), sem_overall = sem(review_overall),
                            sd_aroma = sd(review_aroma), sd_appeareance = sd(review_appearance),sd_overall = sd(review_overall)), 
                        by = beer_style]
question4.means
```

Knowing how each beer style fluctuates on our variables of interest (with our overall score
thrown in for good measure!), let's plot our results.
Note that I have included standard error of the mean error bars as a sanity check to make sure
that each beer's ratings is not going to bleed into the others’ dimensions.
By doing this, we can have a bit more confidence that we are looking at actually has
some meaning.
This plot shows the standard error on each of the two variables we are interested in, and for the
most part it looks as if they are relatively well contained.

```{r}
ggplot(question4.means, aes(x = mean_aroma, y = mean_appearance, color = beer_style)) + 
  geom_point() +
  geom_errorbar(aes(ymin=mean_appearance-sem_appearance, ymax=mean_appearance+sem_appearance), width=.1) +
  geom_errorbarh(aes(xmin=mean_aroma-sem_aroma, xmax=mean_aroma+sem_aroma)) + theme(legend.position="none") +
  labs(title = "Mean Appearance and Aroma", y = "Mean Aroma", x = "Mean Appearance") 
```

This graph has a lot of beers, but what we are interested in is that top right quadrant where both
the average appearance and aroma are maxed out.
Let's zoom in on it and throw in a sizing variable of the overall rating and inspect our graph.

```{r}
ggplot(question4.means[mean_appearance > 4 & mean_aroma > 4], 
       aes(x = mean_aroma, y = mean_appearance, color = beer_style, size = mean_overall)) + 
  geom_point() + xlim(4,4.2) + ylim(4,4.25) +  theme(legend.position="none") +
 # geom_errorbar(aes(ymin=mean_appearance-sem_appearance, ymax=mean_appearance+sem_appearance), width=.1) +
#  geom_errorbarh(aes(xmin=mean_aroma-sem_aroma, xmax=mean_aroma+sem_aroma)) + theme(legend.position="none") +
  labs(title = "Mean Appearance and Aroma", y = "Mean Aroma", x = "Mean Appearance") + 
  geom_text(aes(label=beer_style, hjust = .5, vjust = -.75)) 
```


Looking at this subsection, it appears we have a few choices for beers that score highly on
both Appearance and Aroma.
The American Double / Imperial Stout looks like a good option as it scores higher on how it
looks, but our Russian Imperial Stout has a higher Aroma score.
We could start crunching more numbers here to find "the best" option, or at this point we could
take off our data science hats and put our psychology ones back on (assuming that's what we were wearing...) and run some double-blind experiments on ourselves to make sure that our data is actually lining up with something in the real world.

## Reflections

What started out as a fun weekend project actually turned into a really great learning experience.
I've definitely put a few solid hours into this and have gotten a lot out of it.
I learned that my friends actually know a TON about beer and data science, that [git-lfs](https://git-lfs.github.com/) is something I wish I would have known about earlier,
and that I'm actually looking forward to doing more blogging in the future.

I will probably have to go MIA for the next two weeks since my General Exams are coming up in
early February, but I'm sure I will be back at it come late March.
Until then, all I can hope for is that someone who is looking to hire a junior data scientist over the
summer will come across these posts and think I might be a good temporary addition to
their team.
